---
title: "eda_sean"
author: "Sean Li"
date: "4/22/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exploratory Data Analysis

```{r}
library(tidyverse)
```
Loading in datasets
```{r}
agri_employ <- read_csv("datasets/Agricultural Employment.csv")
agri_data <- read_csv("datasets/Agriculture Dataset.csv")
crops <- read_csv("datasets/Crop Production.csv")
exp_invest <- read_csv("datasets/Expenditure Loans and Investment into Agriculture.csv")
food <- read_csv("datasets/Food Insecurity.csv")
pov_rate <- read_csv("datasets/Poverty Rate.csv")
```



```{r}
povertycountries <- merge(crops,pov_rate, by.x = c('Area','Year'), by.y = c('Country','Year'))
povertycountries <- povertycountries %>% 
  select(Area, Year, Item,`Production - tonnes`, `Percent of Population Below Poverty Line`)
```

```{r}
povertycountries %>% 
  filter(Year == '2020' | Year == '2019' | Year == '2018') %>% 
  pivot_wider(names_from = Item, values_from = `Production - tonnes`) 
  
```


```{r}
top_countries_percrop <- povertycountries %>% 
  filter(Year == '2020' | Year == '2019' | Year == '2018') %>% 
  group_by(Item, Area) %>% 
  summarise(average = mean(`Production - tonnes`)) %>% 
  arrange(desc = TRUE)

```



```{r}
unique(crops$Area)
```

Steps for animated ggplot. The goal is to look at correlation between expenditure and environmental impact? 

Which countries would benefit the most from foreign investment with respect to their development and environmental impact?

X: Total Expenditure on Agriculture, forestry, fishing (2015 US$)
Y: CO2 Emissions Intensity aggregated (CO2eq/kg of product)
Facet: Developing Country or Continent
Size: GDP per capita
Color: Country


1. Merge data 
  - need each row to be country in a specific year
2. Plot
3. Animate


```{r}
#filling NA with 0
agri_data[is.na(agri_data)] = 0 
#getting total emissions aggregated
filted_emissions_dat <- agri_data %>% 
  mutate(total_emissions = `Farm-gate emissions - Emissions (CO2eq) (AR5) - kilotonnes` + 
           `Food Transport - Emissions (CO2eq) (AR5) - kilotonnes` + `Waste - Emissions (CO2eq) (AR5) - kilotonnes` +
           `Cereals excluding rice-Emissions (CO2eq)-kilotonnes` + `Eggs, hen, in shell-Emissions (CO2eq)-kilotonnes` +
           `Meat, buffalo-Emissions (CO2eq)-kilotonnes` + `Meat, cattle-Emissions (CO2eq)-kilotonnes` +
           `Meat, chicken-Emissions (CO2eq)-kilotonnes` + `Meat, goat-Emissions (CO2eq)-kilotonnes` +
           `Meat, pig-Emissions (CO2eq)-kilotonnes` + `Meat, sheep-Emissions (CO2eq)-kilotonnes` + `Milk, whole fresh buffalo-Emissions (CO2eq)-kilotonnes` + `Milk, whole fresh camel-Emissions (CO2eq)-kilotonnes` + `Milk, whole fresh cow-Emissions (CO2eq)-kilotonnes` +
           `Milk, whole fresh goat-Emissions (CO2eq)-kilotonnes` + `Milk, whole fresh sheep-Emissions (CO2eq)-kilotonnes`+
           `Rice, paddy-Emissions (CO2eq)-kilotonnes`) %>% 
  select(Area, Year, Population,`GDP  (2015 US$, mil)`, `Temperature change °C`, total_emissions)

#select important columns from exp_invest
filtered_expenses <- exp_invest %>% 
  mutate(`Agriculture Spending` = `Agriculture, forestry, fishing (General Government) - Value US$, 2015 prices - millions`,
         `Foreign Investment` = `Foreign Investment into Agriculture, Forestry and Fishing - Value US$, 2015 prices - millions`) %>% 
  select(Area,Year,`Agriculture Spending`, `Foreign Investment`)
```

Merge and plot

```{r}
library(gganimate)
library(gifski)
library(png)
```


20K GDP Per Capita
- 14.33 

merging and GDP Per Capita 
```{r}



combined <- merge(filted_emissions_dat,filtered_expenses, by = c("Area","Year")) %>% 
  filter(Year > 1999)

combined <- combined %>% 
  mutate(GDP_Per_Capita = round(`GDP  (2015 US$, mil)`*1e6 / Population, 2)) %>% 
  mutate(developed = ifelse(GDP_Per_Capita*1.1433 > 2*1e4,"Developed","Developing"))
  
```


animation stuff
```{r}

animation <- ggplot(combined, aes(x = `Foreign Investment`, y = total_emissions, size = `GDP  (2015 US$, mil)`, colour = developed)) + 
  geom_point(alpha = 0.5, show.legend = FALSE)+
  scale_fill_gradient() +
  #scale_x_log10(limits = c(50,1e5)) +
  xlim(0,400)+
  scale_y_log10(limits = c(100,1e7)) +
  scale_size(range = c(2, 12)) + 
  facet_wrap(~developed)+
  labs(subtitle = 'Year: {frame_time}', x = 'Agriculture Expenditure in millions (US$)', y = 'CO2 Emissions (Kilotonnes)', title = 'Relationship Between Agriculture Spending and CO2 Emissions') +
  transition_time(as.integer(Year)) +
  #shadow_mark(alpha = 0.3, size = 0.5)+
  ease_aes('linear', interval = 0.001) +
  theme_bw()

animate(animation, end_pause = 3,
         nframes = 100, duration = 20)

```
1. there is move erratic movement with not developed countries
2. less developed countries have a high CO2 emission level compared to total agriculture expenditure
3. generally there is a positive trend between the two.

```{r}
#saving the animation
anim_save("agriculture.gif")
```

```{r}
watered <- combined %>% 
  filter(Area == "Syrian Arab Republic" | Area == "Guatemala" | Area == "South Africa" | Area == "Madagascar" | Area == "Mexico") %>% 
  filter(total_emissions<25000)
```



no relationship between foreign investment and agriculture spending
```{r}
foreign <- ggplot(combined, aes(y = total_emissions,
                       x = `Foreign Investment`, color = developed)) +
  geom_point(alpha = 0.3) +
  geom_point(data = watered, aes(y = total_emissions,
                       x = `Foreign Investment`, color = 'Target')) +
  geom_smooth(method = 'lm', se = FALSE)+
  scale_x_log10(limits = c(1e0,1e4))+
  scale_y_log10(limits = c(1e3,1e7)) +
  labs(x = 'Foreign Investment into Agriculture in millions (US$)', y = 'CO2 Emissions (Kilotonnes)', title = 'Relationship Between Foreign Investment and CO2 Emissions', color = guide_legend(title = "Country Type")) +
  theme_bw()
foreign + scale_color_manual(values = c("#00AFBB", "#e8a30e", "#FC4E07")) 
```

```{r}
ggsave("foreign.png")
```


